---
title: "Deforestation and development in Caatinga"
author: "Lucas Alencar; Luke Parry & Felipe Melo"
date: "25/04/2022"
output: html_document
---

This markdown file was built to develop, share and test the reproducibility of data and analysis from my thesis' chapter one.

<br/>
The first step was classify (and locate) the municipalities of Caatinga in three deforestation frontiers according to their total deforested area. We also calculate the mean annual rate of deforestation inside each frontier group

```{r message=FALSE, warning=FALSE}
library(readxl)
library(here)
library(geobr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(cowplot)
library(tidyr)
library(ggsignif)
library(lme4)
library(emmeans)
library(ggpubr)
library(CBPS)
```

<br/>

Import all data necessary to reproduce the results. Here we are only importing final datatable (and other files) ready for analysis. If you want all scripts made to create the final datatable and other objects, please contatc me at alencar.lucasc@gmail.com
```{r}
read_xlsx(here("data/dbcap1_rma.xlsx"))->dbcap1_rma
read_xlsx(here("data/db2cap1_cbps_clean.xlsx"))-> dbcap1_cbps
```
<br/>

To create the map of figure 1, it is necessary to download some shapes and manipulate the data table
```{r message=FALSE, warning=FALSE}
read_state() %>% 
  filter(abbrev_state %in% c("BA","SE","PE", "AL","PB","RN","CE","PI", "MG"))->uf_caat #Download and filter polygons of brazilian states with Caatinga from oficial data source (IBGE)

read_municipality(simplified = F) ->mun_br #download polygons of brazilian municipalities

```
<br/>

###Map of Caatinga municipalities categorized in three deforestation frontiers and their respective total deforested area and rates of deforestation
```{r}
dbcap1_rma%>%
  select(code_muni, nvc.perc_10, def.stage_10)%>%
  left_join(y = mun_br, by = "code_muni") -> dbcap1_rma.map
  
ggplot()+  
  geom_sf(data=dbcap1_rma.map$geom,
          aes(fill = factor(dbcap1_rma$def.stage_10)), lwd = 0.1, na.rm = T)+
  scale_fill_manual(name = "Deforestation frontier",
                    labels = c("Initial", "Intermediate", "Advanced"),
                    values = c("#5c3811", "#ac6c13", "#ffa600"),
                    na.value = "Grey60")+
  geom_sf(data=uf_caat, fill="transparent", color = "black", lwd = 0.3)+
  coord_sf(xlim = c(-47,-33), ylim = c(-18,-3))+
  geom_text(data = uf_caat, aes(x= c(-42, -39.5,-36.5,-35.5, -34.5, -34.4, -36, -39,-42.4),
                              y = c(-16.8, -15, -11, -10, -8.5, -7, -4.7, -2.9, -5),
                              label = c("MG", "BA", "SE", "AL", "PE", "PB", "RN", "CE", "PI")),
            size = 2)+
  theme_map()+
  theme(legend.position = c(0.65, 0.1),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))-> defStage_map

dbcap1_rma %>% 
  select(code_muni, nvc.perc_10, def.stage_10) %>% 
  glimpse %>% 
  ggplot()+
  geom_boxplot(aes(x = as.factor(def.stage_10), y = 100 - nvc.perc_10, fill = as.factor(def.stage_10)))+
  scale_x_discrete(labels = c("Initial", "Intermediate", "Advanced"))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"))+
  ylab("Deforestation (%) in 2010")+
  theme(axis.title.x = element_blank()
        , panel.background = element_blank()
        , axis.line = element_line(colour = "black")
        , legend.position = "none") -> desmat.front_2010

dbcap1_rma %>% 
  select(code_muni, def.stage_10, tx.desmat.perc_91, tx.desmat.perc_00, tx.desmat.perc_10) %>% 
  mutate(mean.defRate = (tx.desmat.perc_91 + tx.desmat.perc_00 + tx.desmat.perc_10)/3) %>% 
  glimpse %>% 
  ggplot()+
  geom_boxplot(aes(x = as.factor(def.stage_10), y = mean.defRate, fill = as.factor(def.stage_10)))+
  scale_x_discrete(labels = c("Initial", "Intermediate", "Advanced"))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"))+
  ylab("Mean def. rate (%)")+
  xlab("Deforestation frontier (2010)")+
  theme(panel.background = element_blank()
        , axis.line = element_line(colour = "black")
        , legend.position = "none") -> mean_defRate

plot_grid(
  defStage_map,
  plot_grid(
    desmat.front_2010,
    mean_defRate,
    nrow = 2,
    axis = "lr",
    labels = c("b", "c"),
    label_y = c(1, 1.1),
    label_x = c(-0.05,-0.05)
  ),
  rel_widths = c(2, 1),
  labels = "a"
) -> fig1

read_state() -> br_states
read_biomes() %>% 
  filter(name_biome == "Caatinga") ->caat_shp

ggplot()+
  geom_sf(data = caat_shp$geom, color = "transparent", fill = "grey60")+
  geom_sf(data = br_states$geom, fill = "transparent")+
  theme_map() -> inset_map


ggdraw()+
  draw_plot(fig1)+
  draw_plot(inset_map,
            x = -0.03, y = 0.7, width = 0.30, height = 0.30)-> fig1_inset
```

<br/>

##Repeated-measures ANOVA
####In the next step we calculated a rm-ANOVA between groups and across all years. After we make the individual graphics (for each socioeconomic index) and grouped all graphics into a panel (Figure 2)

```{r}
#data prep----
dbcap1_rma%>% 
  select(code_muni, nvc.perc_91, nvc.perc_00, nvc.perc_10, def.stage_91,
         def.stage_2000, def.stage_10, IDHM_E_1991, IDHM_E_2000, IDHM_E_2010,
         IDHM_L_1991, IDHM_L_2000, IDHM_L_2010, IDHM_R_1991, IDHM_R_2000,
         IDHM_R_2010, expov_1991, expov_2000, expov_2010, gini_1991, gini_2000,
         gini_2010, u5mort_1991, u5mort_2000, U5mort_2010)%>% 
  dplyr::rename(nvc.perc_1991 = nvc.perc_91,
         nvc.perc_2000 = nvc.perc_00,
         nvc.perc_2010 = nvc.perc_10,
         def.stage_1991 = def.stage_91,
         def.stage_2010 = def.stage_10,
         u5mort_2010 = U5mort_2010)%>% 
  pivot_longer(cols = -code_muni,
               names_to = c(".value", "year"),
               names_pattern = "(.+)_(.+)",) %>%
  glimpse() -> pl_nvcs

plF_nvcs<- pl_nvcs
plF_nvcs$code_muni<- as.factor(plF_nvcs$code_muni)
plF_nvcs$year<- as.factor(plF_nvcs$year)
plF_nvcs$def.stage<- as.factor(plF_nvcs$def.stage)
glimpse(plF_nvcs)

##IDHM_E----
ggplot(data = plF_nvcs, aes(x = year, y = IDHM_E, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(IDHM_E)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "HDI - Education")+
  geom_signif(y_position = c(0.43, 0.43, 0.43, 0.58, 0.58, 0.58, 0.77, 0.77, 0.77), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","b","b", "c", "d", "d", "e", "e", "f"),
              tip_length = 0)+
  theme_classic(base_size = 14)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())-> rma_idhE

lmer.idhE<- lmer(data = plF_nvcs, formula = IDHM_E ~ def.stage*year + (1 | code_muni))
tab.idhE<- car::Anova(lmer.idhE, type = "II")
summary(lmer.idhE)
pw.idhE<- emmeans(lmer.idhE, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##IDHM_L####
ggplot(data = plF_nvcs, aes(x = year, y = IDHM_L, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(IDHM_L)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontiers")+
  labs(x = "Year", y = "HDI - Longevity")+
  #scale_y_continuous(limits = c(0, 0.85))+
  geom_signif(y_position = c(0.76, 0.76, 0.76, 0.83, 0.83, 0.83, 0.89, 0.89, 0.89), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","a","b", "c", "c", "c", "d", "d", "d"),
              tip_length = 0)+
  theme_classic(base_size = 14)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) -> rma_idhL

lmer.idhL<- lmer(data = plF_nvcs, formula = IDHM_L ~ def.stage*year + (1 | code_muni))
tab.idhL<- car::Anova(lmer.idhL, type = "II")
summary(lmer.idhL)
pw.idhL<- emmeans(lmer.idhL, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##IDHM_R####
ggplot(data = plF_nvcs, aes(x = year, y = IDHM_R, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(IDHM_R)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "HDI - Income")+
  #scale_y_continuous(limits = c(0, 0.85))+
  geom_signif(y_position = c(0.69, 0.69, 0.69, 0.74, 0.74, 0.74, 0.79, 0.79, 0.79), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","b","c", "d", "e", "de", "f", "f", "f"),
              tip_length = 0)+
  theme_classic(base_size = 14)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) -> rma_idhR

lmer.idhR<- lmer(data = plF_nvcs, formula = IDHM_R ~ def.stage*year + (1 | code_muni))
tab.idhR<- car::Anova(lmer.idhR, type = "II")
summary(lmer.idhR)
pw.idhR<- emmeans(lmer.idhR, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##expov####
ggplot(data = plF_nvcs, aes(x = year, y = expov, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(expov)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "Extreme poverty")+
  #scale_y_continuous(limits = c(0, 0.85))+
  geom_signif(y_position = c(96, 96, 96, 76, 76, 76, 51, 51, 51), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","b","c", "d", "e", "de", "f", "g", "fg"),
              tip_length = 0)+
  theme_classic(base_size = 14)-> rma_expov

lmer.expov<- lmer(data = plF_nvcs, formula = expov ~ def.stage*year + (1 | code_muni))
tab.expov<- car::Anova(lmer.expov, type = "II")
summary(lmer.expov)
pw.expov<- emmeans(lmer.expov, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##gini####
ggplot(data = plF_nvcs, aes(x = year, y = gini, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(gini)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "Gini Index")+
  geom_signif(y_position = c(0.94, 0.94, 0.94, 0.84, 0.84, 0.84, 0.81, 0.81, 0.81), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","a","a", "b", "c", "bc", "a", "d", "d"),
              tip_length = 0)+
  theme_classic(base_size = 14) -> rma_gini

lmer.gini<- lmer(data = plF_nvcs, formula = gini ~ def.stage*year + (1 | code_muni))
tab.gini<- car::Anova(lmer.gini, type = "II")
summary(lmer.gini)
pw.gini<- emmeans(lmer.gini, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##u5mort####
ggplot(data = plF_nvcs, aes(x = year, y = u5mort, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(u5mort)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "Under five mortality")+
  geom_signif(y_position = c(155, 155, 155, 110, 110, 110, 55, 55,55), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","a","b", "c", "c", "c", "d", "d", "d"),
              tip_length = 0)+
  theme_classic(base_size = 14) -> rma_u5mort

lmer.u5mort<- lmer(data = plF_nvcs, formula = u5mort ~ def.stage*year + (1 | code_muni))
tab.u5mort<- car::Anova(lmer.u5mort, type = "II")
summary(lmer.u5mort)
pw.u5mort<- emmeans(lmer.u5mort, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

#Figure 2####
ggarrange(rma_idhE, rma_idhL, rma_idhR, rma_expov, rma_gini, rma_u5mort,
          labels = c("a", "b", "c", "d", "e", "f"),
          common.legend = T,
          legend = "bottom"
          )->fig2
```

<br/>

##Next we will do the CBPS analysis
```{r}
##Propensity scores for 2010 model----
dbcap1_cbps %>% 
  mutate(defPerc_2010 = 100 - nvcPerc_2010) %>% 
  glimpse()-> dbcap1_cbps

### Create CBPS----
modelnvc<- CBPS(data = dbcap1_cbps[-c(38,616),], 
              defPerc_2010 ~      
              area_mun +
              p_agro_10 +
              perc_urb_10+
              prodAss_06+
              nascProt_06+
              riosProt_06+
              irrigPerc_06+
              rain_var+
              percProp_S+
              pibAgroPC_2010+
              pibIndPC_2010+
              pibServpubPC_2010+
              carvVeg_10+
              lenha_10,
              method = "exact",
              ATT=0
              )
summary(modelnvc)

### Evaluate CBPS ####
bal.covar<- balance(modelnvc)
bal.covar<- data.frame(original=bal.covar$unweighted, 
                                weighted=bal.covar$balanced)
names(bal.covar)<- c("Pearson r unweighted", "Pearson r IPW")
summary(bal.covar)
boxplot(bal.covar)

##Average Treatment Effect####
set.seed(123)
###outcome IDHM_E----
model.idhmE_2010<- glm(data = dbcap1_cbps[-c(38,616),], 
                       IDHM_E_2010 ~
                         nvcPerc_2010 +
                         I(nvcPerc_2010^2),
                       weights = modelnvc$weights
)
summary(model.idhmE_2010)

lm.morantest(model.idhmE_2010,mat_dist_list, alternative = "two.sided")

sacsarlm(IDHM_E_2010 ~
           nvcPerc_2010 +
           I(nvcPerc_2010^2)+
           area_mun +
           #p_agro_10 +
           perc_urb_10+
           #prodAss_06+
           #nascProt_06+
           riosProt_06+
           irrigPerc_06+
           rain_var+
           percProp_S+
           #pibAgroPC_2010+
           #pibIndPC_2010+
           pibServpubPC_2010,
         #carvVeg_10+
         #lenha_10,
         data = data[-c(38,616),],
         listw = mat_dist_list, 
         type = "sac")->m.idhmE_spat
summary(m.idhmE_spat)
summary(impacts(m.idhmE_spat,
                listw = mat_dist_list,
                R = 500),
        zstats = TRUE)->summ.sac_idhmE

errorsarlm(IDHM_E_2010 ~
             nvcPerc_2010 +
             I(nvcPerc_2010^2)+
             area_mun +
             #p_agro_10 +
             perc_urb_10+
             #prodAss_06+
             #nascProt_06+
             riosProt_06+
             irrigPerc_06+
             rain_var+
             percProp_S+
             #pibAgroPC_2010+
             #pibIndPC_2010+
             pibServpubPC_2010,
           #carvVeg_10+
           #lenha_10,
           data= data[-c(38,616),], listw = mat_dist_list, etype="emixed")->m.idhmE_spat2
summary(m.idhmE_spat)
```

