---
title: "Deforestation and development in Caatinga"
author: "Lucas Alencar; Luke Parry & Felipe Melo"
date: "25/04/2022"
output: html_document
---

This markdown file was built to develop, share and test the reproducibility of data and analysis from my thesis' chapter one.

<br/>
The first step was classify (and locate) the municipalities of Caatinga in three deforestation frontiers according to their total deforested area. We also calculate the mean annual rate of deforestation inside each frontier group

```{r message=FALSE, warning=FALSE}
library(readxl)
library(here)
library(geobr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(cowplot)
library(tidyr)
library(ggsignif)
library(lme4)
library(emmeans)
library(ggpubr)
library(CBPS)
library(spdep)
library(spatialreg)
```

<br/>

Import all data necessary to reproduce the results. Here we are only importing final datatable (and other files) ready for analysis. If you want all scripts made to create the final datatable and other objects, please contatc me at alencar.lucasc@gmail.com
```{r}
read_xlsx(here("data/dbcap1_rma.xlsx"))->dbcap1_rma
read_xlsx(here("data/db2cap1_cbps_clean.xlsx"))-> dbcap1_cbps
```
<br/>

To create the map of figure 1, it is necessary to download some shapes and manipulate the data table
```{r message=FALSE, warning=FALSE}
read_state() %>% 
  filter(abbrev_state %in% c("BA","SE","PE", "AL","PB","RN","CE","PI", "MG"))->uf_caat #Download and filter polygons of brazilian states with Caatinga from oficial data source (IBGE)

read_municipality(simplified = F) ->mun_br #download polygons of brazilian municipalities

```
<br/>

###Map of Caatinga municipalities categorized in three deforestation frontiers and their respective total deforested area and rates of deforestation
```{r}
dbcap1_rma%>%
  dplyr::select(code_muni, nvc.perc_10, def.stage_10)%>%
  left_join(y = mun_br, by = "code_muni") -> dbcap1_rma.map
  
ggplot()+  
  geom_sf(data=dbcap1_rma.map$geom,
          aes(fill = factor(dbcap1_rma$def.stage_10)), lwd = 0.1, na.rm = T)+
  scale_fill_manual(name = "Deforestation frontier",
                    labels = c("Initial", "Intermediate", "Advanced"),
                    values = c("#5c3811", "#ac6c13", "#ffa600"),
                    na.value = "Grey60")+
  geom_sf(data=uf_caat, fill="transparent", color = "black", lwd = 0.3)+
  coord_sf(xlim = c(-47,-33), ylim = c(-18,-3))+
  geom_text(data = uf_caat, aes(x= c(-42, -39.5,-36.5,-35.5, -34.5, -34.4, -36, -39,-42.4),
                              y = c(-16.8, -15, -11, -10, -8.5, -7, -4.7, -2.9, -5),
                              label = c("MG", "BA", "SE", "AL", "PE", "PB", "RN", "CE", "PI")),
            size = 2)+
  theme_map()+
  theme(legend.position = c(0.65, 0.1),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))-> defStage_map

dbcap1_rma %>% 
  dplyr::select(code_muni, nvc.perc_10, def.stage_10) %>% 
  glimpse %>% 
  ggplot()+
  geom_boxplot(aes(x = as.factor(def.stage_10), y = 100 - nvc.perc_10, fill = as.factor(def.stage_10)))+
  scale_x_discrete(labels = c("Initial", "Intermediate", "Advanced"))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"))+
  ylab("Deforestation (%) in 2010")+
  theme(axis.title.x = element_blank()
        , panel.background = element_blank()
        , axis.line = element_line(colour = "black")
        , legend.position = "none") -> desmat.front_2010

dbcap1_rma %>% 
  dplyr::select(code_muni, def.stage_10, tx.desmat.perc_91, tx.desmat.perc_00, tx.desmat.perc_10) %>% 
  mutate(mean.defRate = (tx.desmat.perc_91 + tx.desmat.perc_00 + tx.desmat.perc_10)/3) %>% 
  glimpse %>% 
  ggplot()+
  geom_boxplot(aes(x = as.factor(def.stage_10), y = mean.defRate, fill = as.factor(def.stage_10)))+
  scale_x_discrete(labels = c("Initial", "Intermediate", "Advanced"))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"))+
  ylab("Mean def. rate (%)")+
  xlab("Deforestation frontier (2010)")+
  theme(panel.background = element_blank()
        , axis.line = element_line(colour = "black")
        , legend.position = "none") -> mean_defRate

plot_grid(
  defStage_map,
  plot_grid(
    desmat.front_2010,
    mean_defRate,
    nrow = 2,
    axis = "lr",
    labels = c("b", "c"),
    label_y = c(1, 1.1),
    label_x = c(-0.05,-0.05)
  ),
  rel_widths = c(2, 1),
  labels = "a"
) -> fig1

read_state() -> br_states
read_biomes() %>% 
  filter(name_biome == "Caatinga") ->caat_shp

ggplot()+
  geom_sf(data = caat_shp$geom, color = "transparent", fill = "grey60")+
  geom_sf(data = br_states$geom, fill = "transparent")+
  theme_map() -> inset_map


ggdraw()+
  draw_plot(fig1)+
  draw_plot(inset_map,
            x = -0.03, y = 0.7, width = 0.30, height = 0.30)-> fig1_inset
```

<br/>

##Repeated-measures ANOVA
####In the next step we calculated a rm-ANOVA between groups and across all years. After we make the individual graphics (for each socioeconomic index) and grouped all graphics into a panel (Figure 2)

```{r}
#data prep----
dbcap1_rma%>% 
  dplyr::select(code_muni, nvc.perc_91, nvc.perc_00, nvc.perc_10, def.stage_91,
         def.stage_2000, def.stage_10, IDHM_E_1991, IDHM_E_2000, IDHM_E_2010,
         IDHM_L_1991, IDHM_L_2000, IDHM_L_2010, IDHM_R_1991, IDHM_R_2000,
         IDHM_R_2010, expov_1991, expov_2000, expov_2010, gini_1991, gini_2000,
         gini_2010, u5mort_1991, u5mort_2000, U5mort_2010)%>% 
  dplyr::rename(nvc.perc_1991 = nvc.perc_91,
         nvc.perc_2000 = nvc.perc_00,
         nvc.perc_2010 = nvc.perc_10,
         def.stage_1991 = def.stage_91,
         def.stage_2010 = def.stage_10,
         u5mort_2010 = U5mort_2010)%>% 
  pivot_longer(cols = -code_muni,
               names_to = c(".value", "year"),
               names_pattern = "(.+)_(.+)",) %>%
  glimpse() -> pl_nvcs

plF_nvcs<- pl_nvcs
plF_nvcs$code_muni<- as.factor(plF_nvcs$code_muni)
plF_nvcs$year<- as.factor(plF_nvcs$year)
plF_nvcs$def.stage<- as.factor(plF_nvcs$def.stage)
glimpse(plF_nvcs)

##IDHM_E----
ggplot(data = plF_nvcs, aes(x = year, y = IDHM_E, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(IDHM_E)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "HDI - Education")+
  geom_signif(y_position = c(0.43, 0.43, 0.43, 0.58, 0.58, 0.58, 0.77, 0.77, 0.77), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","b","b", "c", "d", "d", "e", "e", "f"),
              tip_length = 0)+
  theme_classic(base_size = 14)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())-> rma_idhE

lmer.idhE<- lmer(data = plF_nvcs, formula = IDHM_E ~ def.stage*year + (1 | code_muni))
tab.idhE<- car::Anova(lmer.idhE, type = "II")
summary(lmer.idhE)
pw.idhE<- emmeans(lmer.idhE, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##IDHM_L####
ggplot(data = plF_nvcs, aes(x = year, y = IDHM_L, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(IDHM_L)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontiers")+
  labs(x = "Year", y = "HDI - Longevity")+
  #scale_y_continuous(limits = c(0, 0.85))+
  geom_signif(y_position = c(0.76, 0.76, 0.76, 0.83, 0.83, 0.83, 0.89, 0.89, 0.89), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","a","b", "c", "c", "c", "d", "d", "d"),
              tip_length = 0)+
  theme_classic(base_size = 14)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) -> rma_idhL

lmer.idhL<- lmer(data = plF_nvcs, formula = IDHM_L ~ def.stage*year + (1 | code_muni))
tab.idhL<- car::Anova(lmer.idhL, type = "II")
summary(lmer.idhL)
pw.idhL<- emmeans(lmer.idhL, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##IDHM_R####
ggplot(data = plF_nvcs, aes(x = year, y = IDHM_R, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(IDHM_R)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "HDI - Income")+
  #scale_y_continuous(limits = c(0, 0.85))+
  geom_signif(y_position = c(0.69, 0.69, 0.69, 0.74, 0.74, 0.74, 0.79, 0.79, 0.79), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","b","c", "d", "e", "de", "f", "f", "f"),
              tip_length = 0)+
  theme_classic(base_size = 14)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) -> rma_idhR

lmer.idhR<- lmer(data = plF_nvcs, formula = IDHM_R ~ def.stage*year + (1 | code_muni))
tab.idhR<- car::Anova(lmer.idhR, type = "II")
summary(lmer.idhR)
pw.idhR<- emmeans(lmer.idhR, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##expov####
ggplot(data = plF_nvcs, aes(x = year, y = expov, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(expov)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "Extreme poverty")+
  #scale_y_continuous(limits = c(0, 0.85))+
  geom_signif(y_position = c(96, 96, 96, 76, 76, 76, 51, 51, 51), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","b","c", "d", "e", "de", "f", "g", "fg"),
              tip_length = 0)+
  theme_classic(base_size = 14)-> rma_expov

lmer.expov<- lmer(data = plF_nvcs, formula = expov ~ def.stage*year + (1 | code_muni))
tab.expov<- car::Anova(lmer.expov, type = "II")
summary(lmer.expov)
pw.expov<- emmeans(lmer.expov, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##gini####
ggplot(data = plF_nvcs, aes(x = year, y = gini, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(gini)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "Gini Index")+
  geom_signif(y_position = c(0.94, 0.94, 0.94, 0.84, 0.84, 0.84, 0.81, 0.81, 0.81), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","a","a", "b", "c", "bc", "a", "d", "d"),
              tip_length = 0)+
  theme_classic(base_size = 14) -> rma_gini

lmer.gini<- lmer(data = plF_nvcs, formula = gini ~ def.stage*year + (1 | code_muni))
tab.gini<- car::Anova(lmer.gini, type = "II")
summary(lmer.gini)
pw.gini<- emmeans(lmer.gini, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

##u5mort####
ggplot(data = plF_nvcs, aes(x = year, y = u5mort, fill = def.stage)) +
  geom_boxplot(aes(middle = mean(u5mort)))+
  scale_fill_manual(values = c("#5c3811", "#ac6c13", "#ffa600"), labels = c("Initial",
                                                                            "Intermediate",
                                                                            "Advanced"),
                    name = "Deforestation frontier")+
  labs(x = "Year", y = "Under five mortality")+
  geom_signif(y_position = c(155, 155, 155, 110, 110, 110, 55, 55,55), 
              xmin = c(0.7, 0.95, 1.2, 1.7, 1.95, 2.2, 2.7, 2.95, 3.2),
              xmax = c(0.8, 1.05, 1.3, 1.8, 2.05, 2.3, 2.8, 3.05, 3.3),
              annotations = c("a","a","b", "c", "c", "c", "d", "d", "d"),
              tip_length = 0)+
  theme_classic(base_size = 14) -> rma_u5mort

lmer.u5mort<- lmer(data = plF_nvcs, formula = u5mort ~ def.stage*year + (1 | code_muni))
tab.u5mort<- car::Anova(lmer.u5mort, type = "II")
summary(lmer.u5mort)
pw.u5mort<- emmeans(lmer.u5mort, pairwise ~ def.stage*year, pbkrtest.limit = 3621)

#Figure 2####
ggarrange(rma_idhE, rma_idhL, rma_idhR, rma_expov, rma_gini, rma_u5mort,
          labels = c("a", "b", "c", "d", "e", "f"),
          common.legend = T,
          legend = "bottom")
```

<br/>

##Next we will do the CBPS analysis and the SARAR models as robustnesse test

```{r}
##Propensity scores for 2010 model----
### some more data----
dbcap1_cbps %>% 
  mutate(defPerc_2010 = 100 - nvcPerc_2010) %>% 
  glimpse()-> dbcap1_cbps

dbcap1_cbps%>%
  left_join(y = mun_br, by = "code_muni") -> dbcap1_cbps.map

poly2nb(st_as_sf(dbcap1_cbps.map[-c(36, 143, 600, 786),]), queen=TRUE)-> mat_dist #Had to remove 4 municipalities with no neighbours

nb2listw(mat_dist)->mat_dist_list

### Create CBPS----
CBPS(
  data = dbcap1_cbps[-c(36, 143, 600, 786),],
  defPerc_2010 ~
    area_mun +
    p_agro_10 +
    perc_urb_10 +
    prodAss_06 +
    nascProt_06 +
    riosProt_06 +
    irrigPerc_06 +
    rain_var +
    percProp_S +
    pibAgroPC_2010 +
    pibIndPC_2010 +
    pibServpubPC_2010 +
    carvVeg_10 +
    lenha_10,
  method = "exact",
  ATT = 0
) -> modeldef
summary(modeldef)

### Evaluate CBPS ####
bal.covar<- balance(modeldef)
bal.covar<- data.frame(original=bal.covar$unweighted, 
                                weighted=bal.covar$balanced)
names(bal.covar)<- c("Pearson r unweighted", "Pearson r IPW")
summary(bal.covar)
boxplot(bal.covar)

##Average Treatment Effect####
set.seed(123)
###outcome IDHM_E----
model.idhmE_2010<- glm(data = dbcap1_cbps[-c(36, 143, 600, 786),], 
                       IDHM_E_2010 ~
                         defPerc_2010 +
                         I(defPerc_2010^2),
                       weights = modeldef$weights
)
summary(model.idhmE_2010)

lm.morantest(model.idhmE_2010,mat_dist_list, alternative = "two.sided") #confirm the spatial autocorrelation. So we need to test the CBPS model for sensivity to this. We'll use a SARAR (or SAC) models

sacsarlm(IDHM_E_2010 ~
           defPerc_2010 +
           I(defPerc_2010^2)+
           area_mun +
           perc_urb_10+
           riosProt_06+
           irrigPerc_06+
           rain_var+
           percProp_S+
           pibServpubPC_2010,
         data = dbcap1_cbps[-c(36, 143, 600, 786),],
         listw = mat_dist_list, 
         type = "sac")->m.idhmE_spat
summary(m.idhmE_spat)
summary(impacts(m.idhmE_spat,
                listw = mat_dist_list,
                R = 500),
        zstats = TRUE)

###outcome IDHM_L----
model.idhmL_2010<- glm(data = dbcap1_cbps[-c(36, 143, 600, 786),], 
                        IDHM_L_2010 ~
                          defPerc_2010 +
                          I(defPerc_2010^2),
                        weights = modeldef$weights
)
summary(model.idhmL_2010)

lm.morantest(model.idhmL_2010,mat_dist_list, alternative = "two.sided")

sacsarlm(IDHM_L_2010 ~
           defPerc_2010 +
           I(defPerc_2010^2)+
           area_mun +
           perc_urb_10+
           riosProt_06+
           irrigPerc_06+
           rain_var+
           percProp_S+
           pibServpubPC_2010,
         data = dbcap1_cbps[-c(36, 143, 600, 786),],
         listw = mat_dist_list, 
         type = "sac")->m.idhmL_spat
summary(m.idhmL_spat)
summary(impacts(m.idhmL_spat,
                listw = mat_dist_list,
                R = 500),
        zstats = TRUE)

###outcome IDHM_R----
model.idhmR_2010<- glm(data = dbcap1_cbps[-c(36, 143, 600, 786),], 
                    IDHM_R_2010 ~
                      defPerc_2010 +
                      I(defPerc_2010^2),
                    weights = modeldef$weights
)
summary(model.idhmR_2010)

lm.morantest(model.idhmR_2010,mat_dist_list, alternative = "two.sided")

sacsarlm(IDHM_R_2010 ~
           defPerc_2010 +
           I(defPerc_2010^2)+
           area_mun +
           perc_urb_10+
           riosProt_06+
           irrigPerc_06+
           rain_var+
           percProp_S+
           pibServpubPC_2010,
       data = dbcap1_cbps[-c(36, 143, 600, 786),],
       listw = mat_dist_list, 
         type = "sac")->m.idhmR_spat
summary(m.idhmR_spat)
summary(impacts(m.idhmR_spat,
                listw = mat_dist_list,
                R = 500),
        zstats = TRUE)

###outcome expov----

model.expov10<- glm(data = dbcap1_cbps[-c(36, 143, 600, 786),], 
                    expov_2010 ~
                      defPerc_2010 +
                      I(defPerc_2010^2),
                    weights = modeldef$weights
                    )
summary(model.expov10)

lm.morantest(model.expov10, mat_dist_list, alternative = "two.sided")

sacsarlm(expov_2010 ~
           defPerc_2010 +
           I(defPerc_2010^2)+
               area_mun +
               perc_urb_10+
               riosProt_06+
               irrigPerc_06+
               rain_var+
               percProp_S+
               pibServpubPC_2010,
         data = dbcap1_cbps[-c(36, 143, 600, 786),],
         listw = mat_dist_list, 
         type = "sac")->m.expov_spat
summary(m.expov_spat)
summary(impacts(m.expov_spat,
                listw = mat_dist_list,
                R = 500),
        zstats = TRUE)

#outcome gini----

model.gini_2010<- glm(data = dbcap1_cbps[-c(36, 143, 600, 786),], 
                      gini_2010 ~
                        defPerc_2010 +
                        I(defPerc_2010^2),
                      weights = modeldef$weights
)
summary(model.gini_2010)

lm.morantest(model.gini_2010,mat_dist_list, alternative = "two.sided")

sacsarlm(gini_2010 ~
           defPerc_2010 +
           I(defPerc_2010^2)+
           area_mun +
           perc_urb_10+
           riosProt_06+
           irrigPerc_06+
           rain_var+
           percProp_S+
           pibServpubPC_2010,
         data = dbcap1_cbps[-c(36, 143, 600, 786),],
         listw = mat_dist_list, 
         type = "sac")->m.gini_spat
summary(m.gini_spat)
summary(impacts(m.gini_spat,
                listw = mat_dist_list,
                R = 500),
        zstats = TRUE)

#outcome u5mort####
model.u5mort_2010<- glm(data = dbcap1_cbps[-c(36, 143, 600, 786),], 
                      u5mort_2010 ~
                        defPerc_2010 +
                        I(defPerc_2010^2),
                      weights = modeldef$weights
)

summary(model.u5mort_2010)

lm.morantest(model.u5mort_2010,mat_dist_list, alternative = "two.sided")

sacsarlm(u5mort_2010 ~
           defPerc_2010 +
           I(defPerc_2010^2)+
           area_mun +
           perc_urb_10+
           riosProt_06+
           irrigPerc_06+
           rain_var+
           percProp_S+
           pibServpubPC_2010,
         data = dbcap1_cbps[-c(36, 143, 600, 786),],
         listw = mat_dist_list, 
         type = "sac")->m.u5mort_spat
summary(m.u5mort_spat)
summary(impacts(m.u5mort_spat,
                listw = mat_dist_list,
                R = 500),
        zstats = TRUE)
```

<br/>

##Now we will make the graphics of total deforestation area against all socioeconomic index and the final panel

```{r}
#Figures ####
#IDHM_E
ggplot(data = dbcap1_cbps[-c(36, 143, 600, 786),], aes(x = defPerc_2010, y = IDHM_E_2010))+
  geom_point(alpha  = 0.2, color = "#5c3811")+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), lwd = 0.8, fill = "grey80", color = "#ffa600")+
  ylab("HDI - Education")+
  theme_classic()+
  theme(legend.position = "none"
        ,axis.title.x = element_blank()) ->fig.IDHM_E_10

#IDHM_L
ggplot(data = dbcap1_cbps[-c(36, 143, 600, 786),], aes(x = defPerc_2010, y = IDHM_L_2010))+
  geom_point(alpha  = 0.2, color = "#5c3811")+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), lwd = 0.8, fill = "grey80", color = "#ffa600")+
  ylab("HDI - Longevity")+
  theme_classic()+
  theme(legend.position = "none",axis.title.x = element_blank())->fig.IDHM_L_10

#IDHM_R
ggplot(data = dbcap1_cbps[-c(36, 143, 600, 786),], aes(x = defPerc_2010, y = IDHM_R_2010))+
  geom_point(alpha  = 0.2, color = "#5c3811")+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), lwd = 0.8, fill = "grey80", color = "#ffa600")+
  ylab("HDI - Income")+
  theme_classic()+
  theme(legend.position = "none"
        ,axis.title.x = element_blank())->fig.IDHM_R_10

#Extreme poverty
ggplot(data = dbcap1_cbps[-c(36, 143, 600, 786),], aes(x = defPerc_2010, y = expov_2010))+
  geom_point(alpha  = 0.2, color = "#5c3811")+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), lwd = 0.8, fill = "grey80", color = "#ffa600")+
  xlab("Deforestation (%)")+
  ylab("Extreme Poverty")+
  theme_classic()+
  theme(legend.position = "none")->fig.expov_10

#gini
ggplot(data = dbcap1_cbps[-c(36, 143, 600, 786),], aes(x = defPerc_2010, y = gini_2010))+
  geom_point(alpha  = 0.2, color = "#5c3811")+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), lwd = 0.8, fill = "grey80", color = "#ffa600")+
  xlab("Deforestation (%)")+
  ylab("Gini Index")+
  theme_classic()+
  theme(legend.position = "none")->fig.gini_10

#u5mort
ggplot(data = dbcap1_cbps[-c(36, 143, 600, 786),], aes(x = defPerc_2010, y = u5mort_2010))+
  geom_point(alpha  = 0.2, color = "#5c3811")+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), lwd = 0.8, fill = "grey80", color = "#ffa600")+
  xlab("Deforestation (%)")+
  ylab("Under five mortality")+
  theme_classic()+
  theme(legend.position = "none")->fig.u5mort_10

#3figure 3####
ggarrange(fig.IDHM_E_10, fig.IDHM_L_10, fig.IDHM_R_10, fig.expov_10, fig.gini_10, fig.u5mort_10,
          labels = c("a","b", "c", "d", "e", "f"))-> fig3
```

